Ornaments as indicators of social changes in northeastern Taiwan before and after European colonial period

### 2019-12-15 analysis/paper/paper.Rmd
updates the bead name in the plots

```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artifact counts", fig.width=image_width_1_col_inch}
# need to reorder, try https://forcats.tidyverse.org/reference/index.html
# reorder factor for categories (by frequency) and period (manually)
# total number of ornaments for each categories 
ornaments_count <-
  ornaments_period %>% 
  filter(Categories != "Unknown Metal") %>%
  ungroup() %>% # need to upgroup because of rowwise or factor in previous chunk?
  mutate(Categories = fct_lump(Categories, 
                               n = 5)) %>% 
  group_by(Categories) %>%
  count() %>% 
  ungroup() %>% 
  mutate(Categories = fct_reorder(Categories, n))
plot_ornaments_count <- 
ggplot(ornaments_count,
       aes(Categories, n)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(limits = c(0, 160)) +
  labs(y = "Frequency",
       x = "Categories") +
  theme_minimal(base_size = 10)
ggsave(plot = plot_ornaments_count,
       here("analysis", "figures",
       "plot-ornaments-count.jpg"),
       width = 90, 
       height = 40,
       dpi = 300,
       units = "mm")
knitr::include_graphics(here("analysis", 
                             "figures", 
                             "plot-ornaments-count.jpg"))
```

```{r tidy-data-for-table}
# clean some 
ornaments_period_clean <- 
ornaments_period %>% 
  mutate(Type = str_remove(Type, "prism/"),
         Type = ifelse(Type == "small bead", "small (0.5-1 cm)", Type),
         Type = ifelse(Type == "glass bead", "medium (1-2 cm)", Type),
         Type = ifelse(Type == "small I", "plain small", Type),
         Type = ifelse(Type == "small II", "thin small", Type),
         Type = ifelse(Type == "large thin string", "thin large", Type))
ornaments_shape <-
  ornaments_period_clean %>% 
  group_by(period, Categories, Type) %>% 
  tally() %>% 
  spread(period, n, fill = 0) %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring")) %>%
  filter(!Type %in% c("other",
                      "unclassified")) %>% 
  arrange(Categories, desc(`European Presence`))
# making table
knitr::kable(ornaments_shape,
             caption = "Ornament subtype at Kiwulan. The numbers represent artifact counts ")
# some basic stat
sam_post_hole <- nrow(ornaments_period)
total_post_hole <- nrow(kwl_upper)
sample_per <- round(sam_post_hole/total_post_hole*100, 1)

@@

agate_bead_n <- 
  ornaments_count %>% 
  filter(Categories == "Agate bead") %>% 
  pull(n)

@@

metal_bell_n <- 
  ornaments_count %>% 
  filter(Categories == "Bell") %>% 
  pull(n)
```

@@

```{r, plot-counts-by-period, fig.cap = "Frequency of the major ornament across different time periods."}
### delete one plot
# categories of ornaments by period
ornaments_count <-
  ornaments_period %>% 
  filter(!is.na(`6-layers`),
         Categories != "Unknown Metal") %>%
  ungroup() %>% 
  mutate(Categories = fct_lump(Categories, 
                               n = 5)) %>% 
  mutate(Categories = fct_infreq(Categories)) %>% 
  mutate(Categories = fct_rev(Categories)) %>% 
  group_by(period, Categories) %>%
  count()
# histogram facetting for three periods
plot_ornaments_count_three_periods <- 
ggplot(ornaments_count) +
  geom_col(aes(Categories, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(x = "Categories",
       y = "Frequency") +
  theme_minimal(base_size = 10)

@@

```{r agate-bead-fre-plot}
# plot for different shape of agate bead by period
agate_bead_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Agate bead',
         Type != 'other')
base_size_multi_plot <- 8
agate_bead_shape_plot <- 
ggplot(agate_bead_shape) +
  geom_col(aes(reorder(Type, n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "",
       x = "Carnelian bead") +
  theme_minimal(base_size = base_size_multi_plot) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

@@

```{r bell-fre-plot}
# plot for different shape of metal bell by period
bell_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Bell',
         Type != 'unclassified')
bell_shape_plot <-
ggplot(bell_shape) +
  geom_col(aes(reorder(Type,n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "",
       x = "Bell") +
  theme_minimal(base_size = base_size_multi_plot) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

@@

@@

```{r, plot-spatial-pattern-each, fig.cap="Spatial pattern for ornament class by time periods, only those types with more than 5 pieces are shown here"}
ornaments_period_join_selected_types <- 
  ornaments_period_join %>% 
  filter(Categories %in% c('Gold-foil bead', 
                           'Glass bead',
                           'Agate bead'))

>>>>

```{r plot-all-ornaments-count, fig.cap= "Frequency of the major class of ornaments at Kiwulan. Frequency represents artifact counts", fig.width=image_width_1_col_inch}
# need to reorder, try https://forcats.tidyverse.org/reference/index.html
# reorder factor for categories (by frequency) and period (manually)
# total number of ornaments for each categories 
ornaments_count <-
  ornaments_period %>% 
  filter(Categories != "Unknown Metal") %>%
  mutate(Categories = ifelse(Categories == "Agate bead", "Carnelian bead", Categories),
         Categories = ifelse(Categories == "Golden bead", "Gold-foil bead", Categories),
         Categories = ifelse(Categories == "Bell", "Metal bell", Categories)) %>%
  ungroup() %>% # need to upgroup because of rowwise or factor in previous chunk?
  mutate(Categories = fct_lump(Categories, 
                               n = 5)) %>% 
  group_by(Categories) %>%
  count() %>% 
  ungroup() %>% 
  mutate(Categories = fct_reorder(Categories, n))
plot_ornaments_count <- 
ggplot(ornaments_count,
       aes(Categories, n)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(limits = c(0, 160)) +
  labs(y = "Frequency",
       x = "Categories") +
  theme_minimal(base_size = 10)
ggsave(plot = plot_ornaments_count,
       here("analysis", "figures",
       "plot-ornaments-count.jpg"),
       width = 90, 
       height = 40,
       dpi = 300,
       units = "mm")
knitr::include_graphics(here("analysis", 
                             "figures", 
                             "plot-ornaments-count.jpg"))
```

```{r tidy-data-for-table}
# clean some 
ornaments_period_clean <- 
ornaments_period %>% 
  mutate(Type = str_remove(Type, "prism/"),
         Type = ifelse(Type == "small bead", "small (0.5-1 cm)", Type),
         Type = ifelse(Type == "glass bead", "medium (1-2 cm)", Type),
         Type = ifelse(Type == "small I", "plain small", Type),
         Type = ifelse(Type == "small II", "thin small", Type),
         Type = ifelse(Type == "large thin string", "thin large", Type),
         Categories = ifelse(Categories == "Agate bead", "Carnelian bead", Categories),
         Categories = ifelse(Categories == "Golden bead", "Gold-foil bead", Categories),
         Categories = ifelse(Categories == "Bell", "Metal bell", Categories))
ornaments_shape <-
  ornaments_period_clean %>% 
  group_by(period, Categories, Type) %>% 
  tally() %>% 
  spread(period, n, fill = 0) %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring")) %>%
  filter(!Type %in% c("other",
                      "unclassified")) %>% 
  arrange(Categories, desc(`European Presence`))
# making table
knitr::kable(ornaments_shape,
             caption = "Ornament subtype at Kiwulan. The numbers represent artifact counts ")
# some basic stat
sam_post_hole <- nrow(ornaments_period)
total_post_hole <- nrow(kwl_upper)
sample_per <- round(sam_post_hole/total_post_hole*100, 1)

@@

agate_bead_n <- 
  ornaments_count %>% 
  filter(Categories == "Carnelian bead") %>% 
  pull(n)

@@

metal_bell_n <- 
  ornaments_count %>% 
  filter(Categories == "Metal bell") %>% 
  pull(n)
```

@@

```{r, plot-counts-by-period, fig.cap = "Frequency of the major ornament across different time periods."}
### delete one plot
# categories of ornaments by period
ornaments_count <-
  ornaments_period_clean %>% 
  filter(!is.na(`6-layers`),
         Categories != "Unknown Metal") %>%
  ungroup() %>% 
  mutate(Categories = fct_lump(Categories, 
                               n = 5)) %>% 
  mutate(Categories = fct_infreq(Categories)) %>% 
  mutate(Categories = fct_rev(Categories)) %>% 
  group_by(period, Categories) %>%
  count()
# histogram facetting for three periods
plot_ornaments_count_three_periods <- 
ggplot(ornaments_count) +
  geom_col(aes(Categories, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(x = "Categories",
       y = "Frequency") +
  theme_minimal(base_size = 10)

@@

```{r agate-bead-fre-plot}
# plot for different shape of agate bead by period
agate_bead_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Carnelian bead',
         Type != 'other')
base_size_multi_plot <- 8
agate_bead_shape_plot <- 
ggplot(agate_bead_shape) +
  geom_col(aes(reorder(Type, n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "",
       x = "Carnelian bead") +
  theme_minimal(base_size = base_size_multi_plot) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

@@

```{r bell-fre-plot}
# plot for different shape of metal bell by period
bell_shape <-
  ornaments_shape_count %>% 
  filter(Categories == 'Metal bell',
         Type != 'unclassified')
bell_shape_plot <-
ggplot(bell_shape) +
  geom_col(aes(reorder(Type,n), n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "",
       x = " Metal bell") +
  theme_minimal(base_size = base_size_multi_plot) +
  theme(strip.background = element_blank(), 
        strip.text = element_blank())
```

@@

# join two dataset to get coordinate
ornaments_period_join <-
  ornaments_period_clean %>%
  mutate(period = as.factor(period)) %>% # seems need this line first
  group_by(Pit_No, period) %>%
  count(Categories) %>%
  inner_join(AD_data, by = c("Pit_No" = "坑號"))

@@

```{r, plot-spatial-pattern-each, fig.cap="Spatial pattern for ornament class by time periods, only those types with more than 5 pieces are shown here"}
ornaments_period_join_selected_types <- 
  ornaments_period_join %>% 
  filter(Categories %in% c('Gold-foil bead', 
                           'Glass bead',
                           'Carnelian bead'))
